|

### 2023-10-05 chapter_4/src/prr/dcr_color_and_non-political_behaviors_prr.qmd
drafting a little bit more

bibliography: ["../../../assets/neighborhood-chapter-references.bib"]
format:
  hikmah-pdf:
    # Spit out in drafts directory
    latex-output-dir: "../../out/prr"
    # Use biblatex-chicago
    biblatex-chicago: true
    biblio-style: authordate
    biblatexoptions: 
      - backend=biber
      - autolang=hyphen
      - isbn=false
      - uniquename=false
    geometry:
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
      - heightrounded
    linestretch: 2
    fontsize: 12pt
date: today
nocite: |
  @allaire_quarto_2022
execute:
  echo: false
  warning: false
  message: false
  cache: false
params:
  replicate: true
---
{{< pagebreak >}}

@@

In this chapter, I try to accomplish a few goals. First, I want to expand on the previous chapters by examining whether snap-judgments are used to evaluate groups of people. Second, I want to expand on the previous chapters by examining whether the previous findings extend to non-political objects. That is, this chapter seeks to further push my testing of the boundary conditions of the snap-judgment model. The third goal is that this chapter might address an important, but still yet to be solved, puzzle with the literature of partisan sorting that I discuss with more detail in the next section.

@@

Building on this set of evidence, a growing area of the literature suggests that we are not just able to make inferences about the partisanship about other individuals, but we are also able to do this for groups of people through the increasing politicultural distinctions between Republicans and Democrats. @hetherington_prius_2018 demonstrates that Republicans and Democrats are not just different in their policy preferences but rather that they live in different places and have different lifestyles. Others have replicated and built upon this work by suggesting that while there are many different types of lifestyles that Republicans and Democrats can take on, there is increasingly clear distinction between Republicans and Democrat archetypes [@rogers_politicultural_2022]. These politicultural distinctions are clear enough that we train a computer to make accurate guesses (upwards of 80%) of the partisanship by using only images of the cars in a particular neighborhood [@gebru_using_2017].

Though not inherently or necessarily political, our communities often collectively act in political ways. For example, some evidence suggests despite variation between counties' partisan composition and lockdowns discouraging people from leaving their homes during the COVID-19 pandemic, those living in Republican counties spent more time away from home than those living in Democratic counties [@roberts_polarized_2021]. Further, we see a place-based identity for rural individuals that predict partisanship [@cramer_politics_2016; see also @lyons_youre_2021] and characteristics like expression of anti-intellectual attitudes [@lunz_trujillo_rural_2022]. While place-based identities could be separate or informative of partisan identities, some literature argues that partisanship may actually inform which community we choose to make ours.

@@

Some work suggests, however, that these expressed preferences to live near co-partisans do not carryover to migration patterns [see @iyengar_affect_2012]. In fact, there is some evidence suggesting that controlling for everything else but current rates of political sorting, we would have less political homogeneity than we had in 2008 [@martin_does_2020]. Further, evidence by @martin_does_2020 suggests that people switch their party registration after moving as opposed to moving to be more aligned with their party registration.

The literature on partisan sorting seems to conclude that people express a desire to sort into neighborhoods with partisans, but these attitudes do not turn into observable behaviors. The remainder of the chapter provides one explanation for this puzzle by demonstrating how the snap-judgment model may be at play. Much like the literature on the attitude expression of partisan sorting, the snap-judgment model heavily relies on the ways in which seemingly non-political information comes to be associated with one partisan group or another. The model also demonstrates how these snap-judgments are capable of having initial impacts on attitude expression, but its effects weaken in light of explicit and countervailing information.

@@

As discussed thus far in the chapter, non-political objects are often associated with particular partisan leanings. For example, the modal car in a neighborhood conveys the partisan composition of the whole neighborhood [@hetherington_prius_2018;@gebru_using_2017]. In the same spirit as the rest of this book, we do not need to necessarily distinguish between complex forms of information such as whether someone drives a hybrid sedan and a Ford pickup truck. Information giving insight into one's partisanship may even be as subtle as their color choice in car. For a reader who thinks "hypothesizing about car color choice and partisanship is a reach," I agree. As stated at the outset of the chapter, the goal of the chapter is to extend past applications of the snap-judgment model to explicitly partisan objects such as yard signs as we did in [Chapter 2](../../../chapter_2/src/manuscript.qmd) and to really challenge the model by looking at non-political objects that we do not readily associate with partisanship.

@@

Extending the literature on partisanship as a superordinate social identity and affective polarization, we should expect that the partisan composition of a neighborhood would matter to people. In examining yard signs, @makse_politics_2019 demonstrate that people dislike having neighbors that do not share the same partisan affiliation as them. The motivation this comes from a desire to be in the majority of those around you with the same political views [see @van_der_does_strategic_2022]. While this appears to be something important to people, information about the partisan composition of a neighborhood might be inaccessible to most. Therefore, we are motivated to rely on heuristics to help with this [@motyl_how_2020]. One useful and efficient heuristic to convey partisanship is the use of color as other chapters have demonstrated. If we are concerned about the partisanship of our neighbors, we may rely on information such as the color of our potential neighbors' cars to infer their partisanship.

To most of us, we likely do not trust the color of a car as a basis to infer partisanship as other forms of information such as whether a potential neighbor is displaying a political yard sign. However, political yard signs usually only show up every couple of years during the election season and are displayed by those who tend to be highly politically active -- so, they are not universally displayed [@makse_politics_2019]. Additionally, other decisions such as the color of the house are determined by Homeowners Associations (HOAs). The choices of car colors are less constrained by other external factors. So, we may reasonably assume that if we see a neighborhood with a significant number of red cars, we may assume that the local sports team color must be red. However, if we are searching for cues toward the partisan composition of the neighborhood, we may then assume that the neighborhood has a large number of Republicans, whether that is accurate or not. Once we have used that association between the colors red and blue with partisan affiliation, we will form a snap-judgment of our neighbors based on the color of their cars when absent other information. That impression will then elicit a positive or negative attitude toward the neighbors and neighborhood as well as encourage a congruent behavioral reaction such as wanting to avoid the neighborhood or an affinity toward it. This impression and behavioral reaction forms pre-consciously and is quickly adjusted with new information from other sources of information. However, the impression constrains what memories are accessible and therefore constrains the attitude expression and behavioral motivations toward objects even when we collect more information that either fits with or goes against our initial impression.

There are many characteristics of a neighborhood besides car color that can better convey information about the partisanship of a neighborhood such as whether it is mixed use, in an urban or rural setting, whether there are Black Lives Matter flags, and whether the cars tend to be pickup trucks or sedans. However, the role of color is to act as a form of pre-conscious impression formation. While all of these alternative, and better, cues towards partisanship may be present, they are more complicated forms of visual information that will require more laborious processing before attitude formation. The point of the snap-judgment model is to illustrate how extremely simple visual information, such as color, shapes our thinking due to the cognitive processes that underlie impression formation and the effects of impressions after particular neurological paths are made "hot".

@@

The first outcome of interest is the reported feeling thermometer that ranges from 0-100. I divide the response to this outcome variable by 10 and fit an ordered beta regressions [see @kubinec_ordered_2022]. As I am interested in examining how these treatments impact perceptions and resulting preferences based on partisanship, I interact the treatment with the self-reported partisan identity of the participant to examine how a participants' preference for that neighborhood is moderated by the treatment. As I am unable to randomly assign my participants' self-reported partisan identity, I want to ensure that I isolate the main and conditional effects of self-reported partisan identification from confounders such as age, and education. I also want to hold constant potential confounders of the treatment's effect on a participant outside of their partisanship such as sex and colorblindness, as well as attention paid to politics. Using these models will allow me to test the first hypothesis which predicts that people will report liking the neighborhood more when the favorite car color aligns with their partisan group -- red with Republicans and blue with Democrats, as they project the association between the colors red and blue with their partisan group onto the neighborhood. I then examine the second hypothesis which predicts that updating this information with consistent cultural stereotypes of partisan preferences for neighborhood features strengthens one's affect towards the neighborhood as a result of this new information strengthening my participants' certainty about the partisan composition of the neighborhood. These tests do not examine whether it influences one's willingness to move to the neighborhood, only one's preferences toward it. In this way, these tests are much like the outcomes examined in the common experimental tests of the observed positive affect toward neighborhoods with a partisan composition aligned with your own identity. The next outcome of interest examines the degree to which someone is willing to move to the neighborhood on this same information.

@@

if (params$replicate == TRUE) {
  box::use(
    parallel[
      detectCores
    ]
    , brms[
      brmsformula
      , cumulative
      , prior
    ]
    , rstan[stan_model]
    , ./R/helper[
      sim_all
      , discrepancy_df
      , true_positive_hist
    ]
  )
  cores <- detectCores()-1
} else {
  box::use(
   .R/helper[
    true_positive_hist
   ] 
  )
}

@@

>>>>

bibliography: ["../../../assets/neighborhood-chapter-references.bib"]
format:
  hikmah-pdf:
    # Spit out in drafts directory
    latex-output-dir: "../../out/prr"
    # Use biblatex-chicago
    biblatex-chicago: true
    biblio-style: authordate
    biblatexoptions: 
      - backend=biber
      - autolang=hyphen
      - isbn=false
      - uniquename=false
    geometry:
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
      - heightrounded
    linestretch: 2
    fontsize: 12pt
date: today
nocite: |
  @allaire_quarto_2022
execute:
  echo: false
  warning: false
  message: false
  cache: false
params:
  replicate: false
---
{{< pagebreak >}}

@@

In this chapter, I try to accomplish a few goals. First, I want to expand upon the previous chapters by examining whether snap-judgments are used to evaluate groups of people. Second, I want to expand upon the previous chapters by examining whether the previous findings extend to non-political objects. That is, this chapter seeks to further push my testing of the boundary conditions of the snap-judgment model. The third goal is that this chapter might address an important, but still yet to be solved, puzzle with the literature of partisan sorting that I discuss with more detail in the next section.

@@

Building on this set of evidence, a growing area of the literature suggests that we are not just able to make inferences about the partisanship about other individuals, but we are also able to do this for groups of people through the increasing politicultural distinctions between Republicans and Democrats. @hetherington_prius_2018 demonstrates that Republicans and Democrats are not just different in their policy preferences but rather that they live in different places and have different lifestyles. Others have replicated and built upon this work by suggesting that while there are many different types of lifestyles that Republicans and Democrats can take on, there is increasingly clear distinction between Republicans and Democrat archetypes [@rogers_politicultural_2022]. These politicultural distinctions are clear enough that some scholars trained a computer to make accurate inferences (upwards of 80%) of the partisan composition of a neighborhood by using only Google images of the cars parked there [@gebru_using_2017].

Though not inherently or necessarily political, our communities often collectively act in political ways. For example, some evidence suggests despite variation between counties' partisan composition and lockdowns discouraging people from leaving their homes during the COVID-19 pandemic, those living in Republican counties spent more time away from home than those living in Democratic counties [@roberts_polarized_2021]. Further, we see a place-based identity for rural individuals that predict partisanship [@cramer_politics_2016; see also @lyons_youre_2021] and characteristics like expression of anti-intellectual attitudes [@lunz_trujillo_rural_2022] as well as racism among rural Whites and white grievance among urban and suburban Whites on behalf of ruralites [@dawkins_place_2023]. While place-based identities could be separate or informative of partisan identities, some literature argues that partisanship may actually inform which community we choose to make ours.

@@

Some work suggests, however, that these expressed preferences to live near co-partisans do not appear in migration patterns [see @iyengar_affect_2012]. In fact, there is some evidence suggesting that controlling for everything else but current rates of political sorting, we would have less political homogeneity than we had in 2008 [@martin_does_2020]. Further, evidence by @martin_does_2020 suggests that people switch their party registration after moving as opposed to moving to be more aligned with their party registration.

The literature on partisan sorting seems to conclude that people express a desire to sort into neighborhoods with partisans, but these attitudes do not turn into observable behaviors. The remainder of the chapter applies the snap-judgment model to examine this puzzle. Much like the literature on the attitude expression of partisan sorting, the snap-judgment model heavily relies on the ways in which seemingly non-political information comes to be associated with one partisan group or another. The model also demonstrates how these snap-judgments are capable of having initial impacts on attitude expression, but its effects weaken in light of explicit and countervailing information.

@@

As discussed thus far in the chapter, non-political objects are often associated with particular partisan leanings. For example, the modal car in a neighborhood conveys the partisan composition of the whole neighborhood [@hetherington_prius_2018;@gebru_using_2017]. In the same spirit as the rest of this book, we do not need to necessarily distinguish between complex forms of information such as whether someone drives a hybrid sedan and a Ford pickup truck. Information giving insight into one's partisanship may even be as subtle as their color choice in car. For a reader who thinks "hypothesizing about car color choice and partisanship is a reach," I agree. As stated at the outset of the chapter, the goal of the chapter is to extend past applications of the snap-judgment model to explicitly partisan objects such as yard signs as we did in [Chapter 2](../../../chapter_2/src/manuscript.qmd) and to really challenge the model by looking at non-political objects that we may not readily associate with partisanship.

@@

Extending the literature on partisanship as a superordinate social identity and affective polarization, we should expect that the partisan composition of a neighborhood would matter to people. In examining yard signs, @makse_politics_2019 demonstrate that people dislike having neighbors that do not share the same partisan affiliation as them. The motivation this comes from a desire to be in the majority of those around you with the same political views [see @van_der_does_strategic_2022]. While this appears to be something important to people, information about the partisan composition of a neighborhood might be inaccessible to most. Therefore, we are motivated to rely on heuristics to help with this [@motyl_how_2020]. As the other chapters demonstrate, one useful and efficient heuristic to convey partisanship is the use of color. If we are concerned about the partisanship of our neighbors, we may rely on information such as the color of our potential neighbors' cars (or other pieces of their property for that matter) to infer their partisanship.

To most of us, we likely do not trust the color of a car as a basis to infer partisanship as other forms of information such as whether a potential neighbor is displaying a political yard sign. However, political yard signs usually only show up every couple of years during the election season and are displayed by those who tend to be highly politically active -- so, they are not universally displayed [@makse_politics_2019]. Additionally, other decisions such as the color of the house are determined by Homeowners Associations (HOAs). The choices of car colors are less constrained by other external factors. So, we may reasonably assume that if we see a neighborhood with a significant number of red cars, we may assume that the local sports team color must be red. However, in line with predictions from Affective Intelligence Theory, states of uncertainty and anxiety predict information seeking behavior [see @marcus_anxiety_1993]. That is, if we are searching for cues toward the partisan composition of the neighborhood, we may then assume that the neighborhood has a large number of Republicans, whether that is accurate or not. Once we have used that association between the colors red and blue with partisan affiliation, we will form a snap-judgment of our neighbors based on the color of their cars when absent other information. That impression will then elicit a positive or negative attitude toward the neighbors and neighborhood as well as encourage a congruent behavioral reaction such as wanting to avoid the neighborhood or an affinity toward it. This impression and behavioral reaction forms pre-consciously and is quickly adjusted with new information from other sources of information. However, the impression constrains what memories are accessible and therefore constrains the attitude expression and behavioral motivations toward objects even when we collect more information that either fits with or goes against our initial impression.

There are many characteristics of a neighborhood besides car color that can better convey information about the partisanship of a neighborhood such as whether it is mixed use, in an urban or rural setting, whether there are Black Lives Matter flags, and whether the cars tend to be pickup trucks or sedans. However, the role of color is to act as a form of pre-conscious impression formation. While all of these alternative, and perhaps more accurate, cues towards partisanship may be present, they are more complicated forms of visual information that will require more laborious processing before attitude formation. The point of the snap-judgment model is to illustrate how extremely simple visual information, such as color, shapes our thinking due to the cognitive processes that underlie impression formation and the effects of impressions after particular neurological paths are made "hot".

@@

The first outcome of interest is the reported feeling thermometer that ranges from 0-100. I divide the response to this outcome variable by 10 and fit an ordered beta regression [see @kubinec_ordered_2022]. As I am interested in examining how these treatments impact perceptions and resulting preferences based on partisanship, I interact the treatment with the self-reported partisan identity of the participant to examine how a participants' preference for that neighborhood is moderated by the treatment. As I am unable to randomly assign my participants' self-reported partisan identity, I want to ensure that I isolate the main and conditional effects of self-reported partisan identification from confounders such as age, and education. I also want to hold constant potential confounders of the treatment's effect on a participant outside of their partisanship such as sex and colorblindness, as well as attention paid to politics. Using these models will allow me to test the first hypothesis which predicts that people will report liking the neighborhood more when the favorite car color aligns with their partisan group -- red with Republicans and blue with Democrats, as they project the association between the colors red and blue with their partisan group onto the neighborhood. I then examine the second hypothesis which predicts that updating this information with consistent cultural stereotypes of partisan preferences for neighborhood features strengthens one's affect towards the neighborhood as a result of this new information strengthening my participants' certainty about the partisan composition of the neighborhood. These tests do not examine whether it influences one's willingness to move to the neighborhood, only one's preferences toward it. In this way, these tests are much like the outcomes examined in the common experimental tests of the observed positive affect toward neighborhoods with a partisan composition aligned with your own identity. The next outcome of interest examines the degree to which someone is willing to move to the neighborhood on this same information.

@@

if (params$replicate == TRUE) {
  box::use(
    parallel[
      detectCores
    ]
    , brms[
      brmsformula
      , cumulative
      , prior
    ]
    , rstan[stan_model]
    , ./R/helper[
      sim_all
      , discrepancy_df
      , true_positive_hist
    ]
  )
  cores <- detectCores()-1
} else {
  box::use(
   ./R/helper[
    true_positive_hist
    , discrepancy_df
   ] 
  )
}

@@

From these simulations, we can see that the true positive rate for the estimate of interest (the interaction term) holds up across different sample sizes. To be safe, however, I am going to strive to recruit more than 600 participants total which should give me a sufficient sample to work with even once I have to apply exclusion criteria.
